---
title: "mailing lists"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Libraries Used

```{r }
library(dplyr)
library(data.table)
library(ggplot2)
```

## Hilary Clinton mailing lists
For this project we aim to predict a gender based mail text formation using deep learning methods in python.
As for the initial preparation of the data we'll be using R. 

This dataset is devided into three tables we will combine into one full dataset:

[1] Aliases (232 values) - Consists of a row index, mail alias, personId and gender
[2] EmailReceivers (9306 values) - Consists of a row index, mailId and personId
[3] Emails (7945 values) - Consists of 22 features, although we are only interested in just the senderId, mailId, mail subject                            and mail text.


## Read csv's into data.frames
```{r}
aliases <- read.csv("Aliases.csv")
emailReceivers <- read.csv("EmailReceivers.csv")
emails <- read.csv("Emails.csv")
```

## Display alias table
```{r}
head(aliases,50)
```

## Display emailReceivers table
```{r}
head(emailReceivers,50)
```


## select only relevant features from email table
```{r}
emails <- emails %>%
  select(Id,SenderPersonId,ExtractedSubject,ExtractedBodyText)
```

## Display emails table
```{r}
head(emails,50)
```

## Convert data.frames into data.tables
```{r}
aliases <- data.table(aliases)
emailReceivers <- data.table(emailReceivers)
emails <- data.table(emails)
```




## Join aliases and emailReceivers
```{r}
setkey(aliases,PersonId)
setkey(emailReceivers,PersonId)
Result <- merge(emailReceivers,aliases, all.x=TRUE, allow.cartesian=TRUE)
Result <- Result %>%
  select(EmailId,Alias,Gender)
colnames(Result) <- c("emailId","receiver","receiverGender")
head(Result,50)
```


## Join aliases and emailReceivers
```{r}
setkey(emails,SenderPersonId)
Result2 <- merge(emails,aliases,by.x = "SenderPersonId", by.y = "PersonId", all.x=TRUE, allow.cartesian=TRUE)
Result2 <- Result2 %>%
  select(Id.x,Alias,Gender,ExtractedSubject,ExtractedBodyText)
colnames(Result2) <- c("emailId","sender","senderGender","ExtractedSubject","ExtractedBodyText")
head(Result2,50)
```


## Inspect Result and remove NA values after join
```{r}
summary(Result)
Result <- na.omit(Result[, lapply(.SD, function(x) replace(x, which(x==0), NA))])
summary(Result)
```


## Inspect Result2 and remove NA values after join
```{r}
summary(Result2)
Result2 <- na.omit(Result2[, lapply(.SD, function(x) replace(x, which(x==0), NA))])
summary(Result2)
```

## Join both tables into a full dataset of 8709 observations
```{r}
setkey(Result2,emailId)
setkey(Result,emailId)
mailingLists <- merge(Result2,Result, all.x=TRUE, allow.cartesian=TRUE)
```

## Inspect mailingLists and remove NA values after join
```{r}
summary(mailingLists)
mailingLists <- na.omit(mailingLists[, lapply(.SD, function(x) replace(x, which(x==0), NA))])
summary(mailingLists)
```

## remove duplicate values
```{r}
setkey(mailingLists)
m <- unique(mailingLists)
```

## In order to estimate mail text length we'll convert mail text into a string
```{r}
mailingLists$ExtractedBodyText <- as.character(mailingLists$ExtractedBodyText)
```

## mail length feature
```{r}
mailingLists[,mailLength := nchar(ExtractedBodyText)]
str(mailingLists)
```

## Instead of counting words in a text we'll count spaces which is a very close estimate for word count,
Build a function
```{r}
countSpaces <- function(s) { sapply(gregexpr(" ", s), function(p) {sum(p>=0)})}
```

## spaces feature
```{r}
mailingLists <- transform(mailingLists, spaces = countSpaces(mailingLists$ExtractedBodyText))
```

## Look into the dataset
It seems that we have way to much conversations (8580), let's clean the majority of observations
```{r}
summary(mailingLists)
```

## spaces count graph
By looking at this graph and also on the summary and the data itself, most mails are short but there are texts which are extremely long or extremely short
```{r}
ggplot(mailingLists, aes(x = spaces)) +
  geom_histogram(stat = "bin", binwidth=100)
```

## convert to data.frame
```{r}
mailingLists <- data.frame(mailingLists)
```

## Shrinking the dataset
since at the 75th percentile mail texts consist of 27 words we decided to limit texts with above 21 words, and in order to have a bottom limit we chose texts with upto 200 words. 
Now we have 2266 observations, of which 1683 are females and 561 are male senders
```{r}
mailingLists <- mailingLists %>%
  filter(spaces < 200 & spaces > 20)
```


## The next tasks will be performed in Python


